---
title: "Amplify Package: Double seasonal daily time series (Update 0.2.0)"
date: "09/15/2020"
author: Jiaqi (Jackie) Yin
output: 
    github_document:
      html_preview: true
      toc: yes
      toc_depth: 2
      df_print: kable
---

I have leveraged more features to the Amplify Package. Now we release the updated version 0.2.0. The latest one update functions `modify_pred`, `smooth_ts`. `find_adj_date`, and `select_model`. For the first three functions, I have added a mode called `weekday.sep = FALSE`. If the `weekday.sep = TRUE`, one of the seasonal pattern is weekly and time series is implemented respectively according to their weekdays. In `select_model`, I have added cross validation option. See more details below. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 6, cache = TRUE)
```

```{r library, echo=FALSE}
library(tidyverse)
library(forecast)
library(lubridate)
library(tictoc)
```

```{r ggplot-setting, echo = FALSE}
theme_set(theme_bw() + 
            theme(axis.text.x = element_text(color = "grey20", size = 18, face = "plain"),
                  axis.text.y = element_text(color = "grey20", size = 18, face = "plain"),
                  axis.title.x = element_text(color = "grey20", size =20, face = "bold"),
                  axis.title.y = element_text(color = "grey20", size = 20, face = "bold"),
                  legend.text = element_text(size = 18),
                  legend.title = element_text(size = 20, face = "bold"),
                  strip.text.x = element_text(size=18, face="bold")))
```
Our package is for double seasonal daily time seriers. Data features include outliers, variation increasing/decreasing, and double seasonality. We build multiplicative modeling for time series.

In this document, we explain how to use our R package. 

To showing the steps, we forecast the some airline supporting tickets for FY20 as an example.

## Install R package

Install the package from github

```{r install, eval=FALSE}
devtools::install_github("placeboo/amplify", force = TRUE)
```
We start from loading our package `Amplify`.

```{r load-our-package}
library(amplify)
```
We have support ticket from 2014-07-01 to 2018-06-10.
A glimpse of the data:
```{r load-data, echo=FALSE}
load("data/ticket.RData")
tickets %>%
    filter(date < ymd(20140710))
```

In the package, we have a holiday data set. For forecasting, you can add your own holiday information. It should include the date and the name of the holidays. Have a look:

```{r load-holiday}
data("holiday")
head(holiday)
```
```{r, echo= FALSE}
tickets = tickets %>%
  filter(date <= ymd(20190630))
```

## Smoothing 

The whole data set is our training set. We aim to predict the daily tickets from 2019-07-01 to 2020-06-30. First we detect the outiers and impute them with reasonable value.

```{r smooth}
smooth.ls = smooth_ts(data = tickets,
                      vars = c("date", "ticket"),
                      par = list(lambda = 1.5, s = 7),
                      weekday.sep = TRUE)

train.dat = cbind(tickets, smooth.ls$ts.dat)
head(train.dat) 
```

## Validation

We split the data set into subtraining set and validation set to find parameters.
There are 5 parameters $\{\alpha, \beta, \gamma, \omega, \phi \}$ in the model. We use grid search to find a set of optimal parameters. The searching process has been optimized by parallel computing. We also implement cross validation.

```{r split-data}
train.start = 20140701
train.end = 20190630

subtrain.end = 20190430
valid.start = 20190501

valid.ls = train_test_split(train.dat,
                            train.window = c(train.start, subtrain.end),
                            test.window = c(valid.start, train.end))

subtrain.dat = valid.ls$train.dat
valid.dat = valid.ls$test.dat
```

```{r valid}
tic("grid search")
select.ls = select_model(train.y = subtrain.dat$ts,
                         valid.y = valid.dat$ts,
                         grid.search = TRUE,
                         search.length = c(-0.5, 0.5), 
                         length.out = 5,
                         CV = TRUE,
                         kfold = 4)
toc()
```

The list `select.ls` contains the model which has the smallest MAPE for validation set, and the validation results.

```{r cv}
cv = select.ls$cv
head(cv) 
```

We pick the parameters which yield the lowest MAPE. 

```{r mape}
cv[which.min(cv$mape), ]
```

## Prediction

We first build a future data frame

```{r build-future}
pred.dat = build_date(20190701, 20200630)
head(pred.dat) 
```
Within the data frame, we fill the predicted tickets. 
```{r predict}
model = select.ls$model

pred.ls = forecast(model, h = nrow(pred.dat))
pred = as.numeric(pred.ls$mean)
pred.dat = pred.dat %>%
  mutate(pred = pred)

head(pred.dat) 
```

## Holiday Modification

We define the distance between the actual and predictiona and then adjust the holidays.

```{r resid}
tmp = cbind(train.dat, pred = as.numeric(fitted(model)))
tmp = tmp %>%
  mutate(resid = (ticket - pred) / pred)
```

In order to capture floating holidays such as Easter, we add more information about Holidays. 

```{r holiday}
modify.ls = modify_pred(tmp,
                        vars = c("date", "resid"),
                        time.window = c(20190701, 20200630),
                        par = list(lambda = 1.5, n = 3), 
                        holiday = holiday)

h.hat =  modify.ls$h.mat$h.hat
h.hat[h.hat > 0] = 0
pred.dat = pred.dat %>%
  mutate(h.hat = h.hat,
         pred_modify = pred * (1 + h.hat))
```

```{r plot, fig.cap="Prediction of FY20"}
pred.dat %>%
  ggplot(aes(x = date, y = pred_modify)) +
  geom_point(size = 2, alpha = 0.6) + 
  geom_line() +
  scale_x_date(name = "Date",
                date_breaks = "1 month") + 
  ylab("Tickets") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

